---
title: "COD_20240120_HOST_5_phage_QC"
author: "Minsik Kim"
date: "2024-01-20"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
        df_print: paged
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# Loading packages

```{r warning=FALSE, message=FALSE, setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Personal)/Project_SICAS2_microbiome/5_Scripts/MGK/Host_depletion_git"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("readxl", "phyloseq", "tidyverse", "pacman", "yaml", "ggplot2", "vegan", "microbiome", "ggpubr", "viridis", "decontam", "gridExtra", "ggpubr", "lme4", "lmerTest", "writexl", "harrietr", "Maaslin2", "ggtext", "ggpmisc", "gamm4", "reshape2", "kableExtra", "knitr", "ggtree", "car", "mediation", "lemon", "qvalue")
        
YAML_header <-
'---
title: "Host-DNA depletion analysis"
author: "Minsik Kim"
date: "2024.01.20"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230925"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
    .libPaths(path_library)

    require(pacman)
    pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))

    knitr::opts_knit$set(root.dir = path_working)

    str_libraries |> unique() |> sort() -> str_libraries
    pacman::p_load(char = str_libraries)

    set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
    cat("Line Zero Environment:\n\n")
    paste("R:", pacman::p_version(), "\n") |> cat()
    cat("Libraries:\n")
    for (str_libraries in str_libraries) {
        paste(
            "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
            "\n", sep = ""
        ) |> cat()
    }
    paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
    paste("    Library Path:", path_library, "\n") |> cat()
    paste("    Working Path:", path_working, "\n") |> cat()
    paste("Seed:", seed, "\n\n") |> cat()
    cat("YAML Header:\n")
    cat(YAML_header)
}
FUN.LineZero.Report()


```



## Analysis prep {.tabset}

### Loading data

```{r warning=F, message=FALSE, "Data loading"}
phage_example <- read.csv("/Users/minsikkim/Dropbox (Personal)/@minsik/project_ucas/Marker-magu_20240112/P1039_markger_magu_output/p1039.combined_profile.tsv", sep = "\t")
```

# exploratory data analysis 

```{r warning=F, message=FALSE, "Data wrangling"}
phage_example$SGB
phage_example$P_phase4_control
phage_example$marker.magu_genes
phage_example$S_high
phage_example$lineage %>% str_split_fixed(pattern = "[|]", 8) %>% data.frame %>% .[,8] %>% duplicated
phage_example$lineage %>% str_split_fixed(pattern = "[|]", 8) %>% data.frame %>% .[350,] 
phage_example$lineage %>% str_split_fixed(pattern = "[|]", 8) %>% data.frame %>% .[100,] 
phage_example <- read.csv("/Users/minsikkim/Dropbox (Personal)/@minsik/project_ucas/Marker-magu_20240112/P1039_markger_magu_output/p1039.combined_profile.tsv", sep = "\t")
phage_example$SGB %>% str_split_fixed(pattern = "[|]", 2) %>% data.frame %>% .[100,] 
grep("59748", phage_example$SGB)

grep("59748", phage_example$SGB)
```

# Subsetting Viral tax_table & making viral phyloseq object



```{r warning=F, message=FALSE, "Data wrangling"}
phage_example_virus <- subset(phage_example, phage_example$lineage %>% grepl("virus", .))
tax_table_virus <- phage_example_virus$lineage %>% str_split_fixed(pattern = "[|]", 7) %>% {
        data_frame <- data.frame("Kingdom" = .[,1],
                   "Phylum" = .[,2],
                   "Class" = .[,3],
                   "Order" = .[,4],
                   "Family" = .[,5],
                   "Genus" = .[,6],
                   "Species" = .[,7])
        data_frame
}
rownames(tax_table_virus) <- tax_table_virus$Species


row.names(phage_example_virus) <- phage_example_virus$SGB

data_virus <- phage_example_virus%>% select(names(phage_example_virus)[365:373])
data_virus

phage_example_virus <- phage_example_virus %>%
        select(-c(names(phage_example_virus)[365:373])) %>% 
        select(-c("SGB"))
```

# Generating sample_data

the inference about sample type was given by my first impression from the sample_ids (as this is preliminary analysis)

```{r}

sample_data <- data.frame(sample_names = phage_example_virus %>% names) %>%
        mutate(sample_type = ifelse(grepl("mock", sample_names), "Mock DNA",
                                    ifelse(grepl("PosCtrl", sample_names), "Mock community",
                                           ifelse(grepl("dust", sample_names), "Dust samples",
                                                  ifelse(grepl("phase", sample_names), "Ambiguous Ctrls",
                                           "Other samples"))))) %>%
        column_to_rownames("sample_names")
                                    
phage_example_virus
vphyloseq <- merge_phyloseq(sample_data = sample_data(sample_data),
               otu_table = otu_table(phage_example_virus, taxa_are_rows = T),
               tax_table = tax_table(tax_table_virus %>% as.matrix()))


```


# Calculating species richness

```{r}

data <- otu_table(vphyloseq) %>% data.frame
S.obs <- (as.data.frame(otu_table(vphyloseq)) != 0) %>% colSums() %>% data.frame()

#mergind daata - species richenss with sample_data
sample_data(vphyloseq) <- {
        data <- sample_data(vphyloseq)        
        data$species_richness <- S.obs$.
        sample_data(data)
        data
}



sample_data(vphyloseq)

```

# plotting species richenss (viral)

```{r}

sample_names(vphyloseq)
sample_data(vphyloseq) %>%
        as.data.frame() %>%
        ggplot(., aes(x = sample_type, y = species_richness, fill = sample_type)) +
        geom_boxplot() +
        theme_classic(base_size = 10) +
        scale_fill_brewer(type = "qual", palette = 6) +
        ylab("Viral species richness") +
        #theme(axis.text.x = element_text(angle=20, ))
        xlab("Sample type") 


```

```{r}


summary_dataframe <- sample_data(vphyloseq) %>% data.frame %>% group_by(sample_type) %>% summarise(`Species richness` = paste(round(mean(species_richness), 2), "±", round(sd(species_richness), 2))) %>% data.frame()


```
# Marker-Magu data processing

## Laoding Marker-MAGu data

```{r}
marker_magu_reads1 <- read.csv("/Users/minsikkim/Dropbox (Personal)/Project_SICAS2_microbiome/4_Data/1_Raw/Sequencing_Baylor_Metagenomics/20240209_PQ00331_marker_magu_output/PQ00331.AlignedReads.tsv", sep = "\t", check.names = F)

marker_magu_tax1 <- read.csv("/Users/minsikkim/Dropbox (Personal)/Project_SICAS2_microbiome/4_Data/1_Raw/Sequencing_Baylor_Metagenomics/20240209_PQ00331_marker_magu_output/PQ00331.combined_profile.tsv", check.names = F)

marker_magu_reads2 <- read.csv("/Users/minsikkim/Dropbox (Personal)/Project_SICAS2_microbiome/4_Data/1_Raw/Sequencing_Baylor_Metagenomics/20240209_PQ00430_marker_magu_output/PQ00430.without.spikeins.AlignedReads.tsv", sep = "\t", check.names = F)

marker_magu_tax2 <- read.csv("/Users/minsikkim/Dropbox (Personal)/Project_SICAS2_microbiome/4_Data/1_Raw/Sequencing_Baylor_Metagenomics/20240209_PQ00430_marker_magu_output/PQ00430.without.spikeins.combined_profile.tsv", check.names = F)

```

## Loading phyloseq data

Loading phylose data to extract sample metadata

```{r}
phyloseq_unfiltered <- readRDS("/Users/minsikkim/Dropbox (Personal)//Project_SICAS2_microbiome/4_Data/2_Tidy/Phyloseq/PHY_20230521_MGK_host_tidy.rds")

sample_data <- phyloseq_unfiltered$phyloseq_count %>% sample_data


sample_data$baylor_id
```


# Bibliography

```{r warning=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
    str_libraries |>
        pacman::p_citation() |>
        print(bibtex = FALSE) |>
        capture.output() %>%
        .[-1:-3] %>% .[. != ""] |>
        stringr::str_squish() |>
        stringr::str_replace("_", "") |>
        cat()
    cat("\n")
}
#===============================================================================
```
