---
title: "COD_20230604_HOST_5_controls"
author: "Minsik Kim"
date: "2023-06-04"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
        df_print: paged
editor_options: 
  chunk_output_type: inline
  markdown: 
    wrap: 72
---

# Loading packages

```{r warning=FALSE, message=FALSE, setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
knitr::opts_chunk$set(message=FALSE, warning = FALSE)

path_working <- "/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/5_Scripts/MGK/Host_depletion_git/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c(
    "readxl", "phyloseq", "tidyverse", "pacman", "yaml"
)

path_working <- "/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/5_Scripts/MGK/Host_depletion_git"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("readxl", "phyloseq", "tidyverse", "pacman", "yaml", "ggplot2", "vegan", "microbiome", "ggtext")
        
YAML_header <-
'---
title: "COD_20230604_HOST_5_controls"
author: "Minsik Kim"
date: "2032.06.04"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230604"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
    .libPaths(path_library)

    require(pacman)
    pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))

    knitr::opts_knit$set(root.dir = path_working)

    str_libraries |> unique() |> sort() -> str_libraries
    pacman::p_load(char = str_libraries)

    set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
    cat("Line Zero Environment:\n\n")
    paste("R:", pacman::p_version(), "\n") |> cat()
    cat("Libraries:\n")
    for (str_libraries in str_libraries) {
        paste(
            "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
            "\n", sep = ""
        ) |> cat()
    }
    paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
    paste("    Library Path:", path_library, "\n") |> cat()
    paste("    Working Path:", path_working, "\n") |> cat()
    paste("Seed:", seed, "\n\n") |> cat()
    cat("YAML Header:\n")
    cat(YAML_header)
}
FUN.LineZero.Report()


```

# Loading data

## Analysis prep {.tabset}

### Controls of HOST project

**Sequencing result phyloseq object was loaded**

```{r warning=F, message=FALSE, "Data loading"}
# Loading files -----------------------------------------------------------
#loading tidy phyloseq object
phyloseq <- readr::read_rds("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/4_Data/2_Tidy/Phyloseq/PHY_20221129_MGK_host_tidy_tax.rds")


#sample data loading
sample_data <- sample_data(phyloseq$phyloseq_count)



```


### SICAS2 positive controls

**Mock community data from other project / having different lot number was loaded**


``` {r 0.2, warning = FALSE}

phyloseq_control_sample <- read.csv("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/1_Daily_Log/MGK/LOG_20220926_MGK_extraction_data.csv")
phyloseq_control_sample <- subset(phyloseq_control_sample, phyloseq_control_sample$spikein_type != "high" | is.na(phyloseq_control_sample$spikein_type))
#Making metadata
phyloseq_control_sample <- select(phyloseq_control_sample, c("BaylorID", "Type", "extraction_date"))
names(phyloseq_control_sample) <- c("baylor_id", "reagent", "extraction_date")
rownames(phyloseq_control_sample) <- phyloseq_control_sample$biom_id
phyloseq_control_sample$sample_type <- ifelse(phyloseq_control_sample$reagent == "reagent", "Neg.", "Mock")
phyloseq_control_sample$lot_number <- ifelse(phyloseq_control_sample$reagent == "reagent", NA, str_split_fixed(phyloseq_control_sample$reagent, pattern = "_", 4) %>% data.frame %>% .$X4)

phyloseq_control_sample$extraction_date <- as.Date(phyloseq_control_sample$extraction_date, format = c("%m/%d/%y"))
phyloseq_control_sample$lai_id <- paste(phyloseq_control_sample$control, as.Date(phyloseq_control_sample$extraction_date), sep = "_")

#Adding metadata - knead data output
kneaddata_1 <- read.csv("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20220628_PQ00254_SICAS2/P1488_PQ00254_Deliverables_With_Spikeins/ReadStats_w_spikeins.tsv", sep = "\t")
kneaddata_2 <- read.csv("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20220216_Baylor_SICAS2_Biobakery3_All_Kingdom_Corrected_ID/Deliverables/ReadStats.txt", sep = "\t")
kneaddata_1 <- kneaddata_1 %>% select(c(names(kneaddata_2)))

kneaddata_sicas <- rbind(kneaddata_1, kneaddata_2) %>% rename(baylor_id = "SampleID")
phyloseq_control_sample <- merge(kneaddata_sicas, phyloseq_control_sample, by = "baylor_id")

seq_data <- biomformat::read_biom("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20220216_Baylor_SICAS2_Biobakery3_All_Kingdom_Corrected_ID/Deliverables/metaphlan3/metaphlan3.allkingdoms.EstCount.biom")
seq_data2 <- biomformat::read_biom("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20220628_PQ00254_SICAS2/P1488_PQ00254_Deliverables_With_Spikeins/metaphlan3/metaphlan3.allkingdoms.EstCount.biom")


seq_data <- import_biom(seq_data)
seq_data2 <- import_biom(seq_data2)
sample_names(seq_data) <- str_split_fixed(sample_names(seq_data), "-", 2) %>% data.frame() %>% .[,1]
row.names(phyloseq_control_sample) <- phyloseq_control_sample$baylor_id
ctrl_data <- merge_phyloseq(seq_data, sample_data(phyloseq_control_sample))
ctrl_data2 <- merge_phyloseq(seq_data2, sample_data(phyloseq_control_sample))
phyloseq$sicas2_control <- merge_phyloseq(ctrl_data2, ctrl_data)



colnames(tax_table(phyloseq$sicas2_control)) <- c("Domain", "Phylum", "Class", "Order", "Family",  "Genus", "Species")

sample_data(phyloseq$phyloseq_count)$lot_number <- ifelse(sample_data(phyloseq$phyloseq_count)$sample_type == "Mock",
                                                          ifelse(sample_data(phyloseq$phyloseq_count)$extraction_date < "20221212",
                                                          190633,
                                                          213683),
                                                          NA)
phyloseq$positives <- merge_phyloseq(phyloseq$phyloseq_count, phyloseq$sicas2_control) %>% subset_samples(sample_type == "Mock")



```

### Alpha diversity indices

**Alpha diversity indices were calculated and added to the metadata**

 - to eliminate samples with zero OTUs

```{r}           



alpha_diversity <- function(data) {
        otu_table <- otu_table(data) %>% .[colSums(.) !=0]
        S.obs <- rowSums(t(otu_table) != 0)
        sample_data <- sample_data(data)
        data_evenness <- vegan::diversity(t(otu_table)) / log(vegan::specnumber(t(otu_table))) # calculate evenness index using vegan package
        data_shannon <- vegan::diversity(t(otu_table), index = "shannon") # calculate Shannon index using vegan package
        data_hill <- exp(data_shannon)                           # calculate Hills index
        data_dominance <- microbiome::dominance(otu_table, index = "all", rank = 1, aggregate = TRUE) # dominance (Berger-Parker index), etc.
        data_invsimpson <- vegan::diversity(t(otu_table), index = "invsimpson")                          # calculate Shannon index using vegan package
        alpha_diversity <- cbind(S.obs, data_shannon, data_hill, data_invsimpson, data_evenness,data_dominance) # combine all indices in one data table
        sample_data <- merge(data.frame(sample_data), alpha_diversity, by = 0, all = T) %>% column_to_rownames(var = "Row.names")
}
phyloseq_unfiltered <- phyloseq
sample_data(phyloseq_unfiltered$phyloseq_rel) <- sample_data(alpha_diversity(phyloseq$phyloseq_count))
sample_data(phyloseq_unfiltered$phyloseq_count) <- sample_data(alpha_diversity(phyloseq$phyloseq_count)) 
sample_data(phyloseq_unfiltered$phyloseq_path_rpkm) <- sample_data(alpha_diversity(phyloseq$phyloseq_path_rpkm))  
sample_data(phyloseq_unfiltered$positives) <- sample_data(alpha_diversity(phyloseq$positives))  
phyloseq_control_rel <- subset_samples(phyloseq_unfiltered$phyloseq_rel, sample_type == "Mock" | sample_type == "Neg.") #adding data of controls
phyloseq_mock_rel <- transform_sample_counts(phyloseq_unfiltered$positives, function(x) {x/sum(x)})

```

### {-}

# Exploratory analysis

## {.tabset}

### Bar plot of controls (Positive)

**Some possible contaminants were identified in extraction controls**

```{r, warning=FALSE}

#Manipulating phyloseq - only top 10 

phyloseq_control_rel %>% 
        subset_samples(., sample_type == "Mock") %>% 
        tax_table() %>%
        cbind(species20 = "[Others]") %>%
        {top20species <- head(taxa_sums(subset_samples(phyloseq_control_rel,sample_type == "Mock" & S.obs != 0)) %>%
                                data.frame %>%
                                arrange(-.) %>%
                                row.names(), 10)
   .[top20species, "species20"] <- as.character(.[top20species, "Species"])
   .[, 9] <- .[, 9] %>% gsub("s__", "", .) %>% gsub("_", " ", .) %>% paste("<i>", ., "</i>", sep = "")
   phyloseq_temp <- subset_samples(phyloseq_control_rel,sample_type == "Mock" & S.obs != 0)
   tax_table(phyloseq_temp) <- tax_table(.) 
   phyloseq_temp
  } %>%
        plot_bar(., fill="species20") + 
        ylab("Relative abundancne") +
        theme_classic(base_size = 11, base_family = "serif") +
        ggtitle("Postive controls") +
        theme(legend.text = element_markdown()) +
        guides(fill=guide_legend(title="Top 10 species")) +
        facet_wrap (~ factor(treatment, levels = c("Untreated", "lyPMA", "Benzonase", "Host zero", "Molysis", "QIAamp")),
                    scales= "free_x", nrow=1)

#there could be opportunistic pathogens...

                                

```

### Positives of other projects

**Mock community controls of other projects from other sequencing batches / different lot numbers**

Lot# 213683 & 190633 were used for the host depletion study.


The top 10 species are aligning with expected taxa in the mock community besides *Bacillus intestinallis*.

**It seems like Metaphlan failed to assign taxa precisely.**

Expected taxa in Zymo D6300 are

"Pseudomonas_aeruginosa",
"Escherichia_coli",
"Salmonella_enterica",
"Lactobacillus_fermentum",
"Enterococcus_faecalis",
"Staphylococcus_aureus",
"Listeria_monocytogenes",
**"Bacillus_subtilis",**
"Saccharomyces_cerevisiae",
"Cryptococcus_neoformans"

```{r, warning=FALSE}

#Manipulating phyloseq - only top 10 


phyloseq_mock_rel_untreated <- phyloseq_mock_rel %>% 
        subset_samples(treated == 0 | is.na(treated)) 

phyloseq_mock_rel_untreated %>% 
        tax_table() %>%
        cbind(species20 = "[Others]") %>%
        {top20species <- head(taxa_sums(subset_samples(phyloseq_mock_rel_untreated,sample_type == "Mock" & S.obs != 0)) %>%
                                data.frame %>%
                                arrange(-.) %>%
                                row.names(), 10)
   .[top20species, "species20"] <- as.character(.[top20species, "Species"])
   .[, 9] <- .[, 9] %>% gsub("s__", "", .) %>% gsub("_", " ", .) %>% paste("<i>", ., "</i>", sep = "")
   phyloseq_temp <- subset_samples(phyloseq_mock_rel_untreated,sample_type == "Mock" & S.obs != 0)
   tax_table(phyloseq_temp) <- tax_table(.) 
   phyloseq_temp
  } %>%
        plot_bar(., fill="species20") + 
        ylab("Relative abundancne") +
        theme_classic(base_size = 11, base_family = "serif") +
        ggtitle("Postive controls facetted by lot number") +
        theme(legend.text = element_markdown(), axis.text.x = element_blank()) +
        guides(fill=guide_legend(title="Top 10 species")) +
        facet_wrap(~lot_number, scales = "free_x", ncol = 4)
        

```
### {-}

## Focusing on unexpected taxa {.tabset}

### Unexpected species in mock community controls 

**(after removing expected taxa)**

**Significant contaminants were identified - in one sample**

**Others had some unexpected, evenly distributed taxa**

**This could be 1) background contamination or 2) cross-contamination from kingfisher 3) sequencing artifacts.**

```{r, warning=FALSE}
#Manipulating phyloseq - only top 10 





phyloseq_control_rel_contam <- subset_taxa(phyloseq_control_rel , !(taxa_names(phyloseq_control_rel) %in% c("Pseudomonas_aeruginosa_group",
                                              "Escherichia_coli",
                                              "Salmonella_enterica",
                                              "Lactobacillus_fermentum",
                                              "Enterococcus_faecalis",
                                              "Staphylococcus_aureus",
                                              "Listeria_monocytogenes",
                                              "Bacillus_intestinalis",
                                              "Saccharomyces_cerevisiae",
                                              "Cryptococcus_neoformans")
                                              )
                                          )

phyloseq_control_rel_contam <- subset_taxa(phyloseq_control_rel_contam, taxa_sums(phyloseq_control_rel_contam) != 0)
phyloseq_control_rel_contam <- subset_samples(phyloseq_control_rel_contam, sample_type != "Neg." & S.obs != 0)


tax_table(phyloseq_control_rel_contam) %>%
        cbind(species20 = "[Others]") %>%
        {top20species <- head(taxa_sums(phyloseq_control_rel_contam) %>%
                                data.frame %>%
                                arrange(-.) %>%
                                row.names(), 10)
   .[top20species, "species20"] <- as.character(.[top20species, "Species"])
   .[, 9] <- .[, 9] %>% gsub("s__", "", .) %>% gsub("_", " ", .) %>% paste("<i>", ., "</i>", sep = "")
   phyloseq_temp <- phyloseq_control_rel_contam
   tax_table(phyloseq_temp) <- tax_table(.) 
   phyloseq_temp
  } %>%
        plot_bar(., fill="species20") + 
        ylab("Relative abundancne") +
        theme_classic(base_size = 11, base_family = "serif") +
        ggtitle("Contaminants in Zymo mock") +
        theme(legend.text = element_markdown()) +
        guides(fill=guide_legend(title="Top 10 species")) +
        facet_wrap (~ factor(treatment, levels = c("Untreated", "lyPMA", "Benzonase", "Host zero", "Molysis", "QIAamp")),
                    scales= "free_x", nrow=1)


```

### Unexpected taxa from other projects

**Unexpected taxa are quite similar, even from different sequencing batches and different lot# of Zymo D6300**

```{r, warning=FALSE}
#Manipulating phyloseq - only top 10 



phyloseq_mock_rel_untreated_contam <- subset_taxa(phyloseq_mock_rel_untreated , !(taxa_names(phyloseq_mock_rel_untreated) %in% c("Pseudomonas_aeruginosa_group",
                                              "Escherichia_coli",
                                              "Salmonella_enterica",
                                              "Lactobacillus_fermentum",
                                              "Enterococcus_faecalis",
                                              "Staphylococcus_aureus",
                                              "Listeria_monocytogenes",
                                              "Bacillus_intestinalis",
                                              "Saccharomyces_cerevisiae",
                                              "Cryptococcus_neoformans")
                                              )
                                          )

phyloseq_mock_rel_untreated_contam <- subset_taxa(phyloseq_mock_rel_untreated_contam, taxa_sums(phyloseq_mock_rel_untreated_contam) != 0)
phyloseq_mock_rel_untreated_contam <- subset_samples(phyloseq_mock_rel_untreated_contam, sample_type != "Neg." & S.obs != 0)


tax_table(phyloseq_mock_rel_untreated_contam) %>%
        cbind(species20 = "[Others]") %>%
        {top20species <- head(taxa_sums(phyloseq_mock_rel_untreated_contam) %>%
                                data.frame %>%
                                arrange(-.) %>%
                                row.names(), 10)
   .[top20species, "species20"] <- as.character(.[top20species, "Species"])
   .[, 9] <- .[, 9] %>% gsub("s__", "", .) %>% gsub("_", " ", .) %>% paste("<i>", ., "</i>", sep = "")
   phyloseq_temp <- phyloseq_mock_rel_untreated_contam
   tax_table(phyloseq_temp) <- tax_table(.) 
   phyloseq_temp
  } %>%
        plot_bar(., fill="species20") + 
        ylab("Relative abundancne") +
        theme_classic(base_size = 11, base_family = "serif") +
        ggtitle("Zymo mock community - after subtracting expected taxa (by lot number)") +
        theme(legend.text = element_markdown()) +
        guides(fill=guide_legend(title="Top 10 species")) +
        facet_wrap (~lot_number,scales= "free_x", nrow=1)


```
### {-}

## Genus level analysis {.tabset}

**As some unexpected taxa are from the same genus of expected taxa, the same analysis was conducted at genus level**

### Genus level - whole taxa

**Bar plot at genus level showed top 10 genus are expected taxa**

```{r, warning=FALSE}

phyloseq_mock_genus <- tax_glom(phyloseq_mock_rel_untreated, taxrank="Genus")

#Manipulating phyloseq - only top 10 
phyloseq_mock_genus %>% 
        tax_table() %>%
        cbind(species20 = "[Others]") %>%
        {top20species <- head(taxa_sums(subset_samples(phyloseq_mock_genus,sample_type == "Mock" & S.obs != 0)) %>%
                                data.frame %>%
                                arrange(-.) %>%
                                row.names(), 10)
   .[top20species, "species20"] <- as.character(.[top20species, "Genus"])
   .[, 9] <- .[, 9] %>% gsub("g__", "", .) %>% gsub("_", " ", .) %>% paste("<i>", ., "</i>", sep = "")
   phyloseq_temp <- subset_samples(phyloseq_mock_genus,sample_type == "Mock" & S.obs != 0)
   tax_table(phyloseq_temp) <- tax_table(.) 
   phyloseq_temp
  } %>%
        plot_bar(., fill="species20") + 
        ylab("Relative abundancne") +
        theme_classic(base_size = 11, base_family = "serif") +
        ggtitle("Postive controls at genus level") +
        theme(legend.text = element_markdown()) +
        guides(fill=guide_legend(title="Top 10 genus")) 
        
#there could be opportunistic pathogens...

```

### Unexpected genus (after removing expected taxa)

**1. Unexpected genus was 1/10 of unexpected taxa**

**90% of unexpected species seem to be misclassified species**

**other 10% could be contaminants, but again they are quite similar across all the sequencing batch/lot#. **

```{r}

phyloseq_mock_genus_contam <- subset_taxa(phyloseq_mock_genus , !(taxa_names(phyloseq_mock_genus) %in% c("Pseudomonas_aeruginosa_group",
                                              "Escherichia_coli",
                                              "Salmonella_enterica",
                                              "Lactobacillus_fermentum",
                                              "Enterococcus_faecalis",
                                              "Staphylococcus_aureus",
                                              "Listeria_monocytogenes",
                                              "Bacillus_intestinalis",
                                              "Saccharomyces_cerevisiae",
                                              "Cryptococcus_neoformans")
                                              )
                                          )

phyloseq_mock_genus_contam <- subset_taxa(phyloseq_mock_genus_contam, taxa_sums(phyloseq_mock_genus_contam) != 0)
phyloseq_mock_genus_contam <- subset_samples(phyloseq_mock_genus_contam, sample_type != "Neg." & S.obs != 0)

tax_table(phyloseq_mock_genus_contam) %>%
        cbind(species20 = "[Others]") %>%
        {top20species <- head(taxa_sums(phyloseq_mock_genus_contam) %>%
                                data.frame %>%
                                arrange(-.) %>%
                                row.names(), 10)
   .[top20species, "species20"] <- as.character(.[top20species, "Genus"])
   .[, 9] <- .[, 9] %>% gsub("g__", "", .) %>% gsub("_", " ", .) %>% paste("<i>", ., "</i>", sep = "")
   phyloseq_temp <- phyloseq_mock_genus_contam
   tax_table(phyloseq_temp) <- tax_table(.) 
   phyloseq_temp
  } %>%
        plot_bar(., fill="species20") + 
        ylab("Relative abundancne") +
        theme_classic(base_size = 11, base_family = "serif") +
        ggtitle("Zymo mock community - after subtracting expected taxa (by lot number)") +
        theme(legend.text = element_markdown()) +
        guides(fill=guide_legend(title="Top 10 genus")) +
        facet_wrap (~lot_number,scales= "free_x", nrow=1)


```
### {-}

# Results

**1. Some unexpected taxa were identified**

**2. MetaPhlAn clearly misclassified Bacillus subtilis**

**3. There was clear contamination in one sample (lyPMA treated)**

**4. 90% of unexpected were from the same genus of the expected group**

- i.e., unexpected taxa were 3% in relative abundance at the species level but decreased to 0.3% when the data were aggregated to the genus level.

- These taxa were present regardless of *sequencing batch* or *mock community reagent lot#*

**Therefore, it seems most of the unexpected taxa in the mock community are sequencing artifacts**

**Further analysis (using another pipeline) is ongoing**

**Temporarily, genus level analysis can be conducted**

# Bibliography

```{r warning=FALSE}
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
    str_libraries |>
        pacman::p_citation() |>
        print(bibtex = FALSE) |>
        capture.output() %>%
        .[-1:-3] %>% .[. != ""] |>
        stringr::str_squish() |>
        stringr::str_replace("_", "") |>
        cat()
    cat("\n")
}
#===============================================================================
```
