---
title: "COD_20230412_HOST_1_Wrangle"
author: "Minsik Kim"
date: "2023-04-12"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
        df_print: paged
editor_options: 
  chunk_output_type: inline
---


## Loading packages

```{r setup}
#===============================================================================
#BTC.LineZero.Header.1.1.0
#===============================================================================
#R Markdown environment setup and reporting utility.
#===============================================================================
#RLB.Dependencies:
#   knitr, magrittr, pacman, rio, rmarkdown, rmdformats, tibble, yaml
#===============================================================================
#Input for document parameters, libraries, file paths, and options.
#=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=+=
path_working <- "/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c(
    "readxl", "phyloseq", "tidyverse", "pacman", "yaml"
)

path_working <- "/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/5_Scripts/MGK/Host_depletion_git"
path_library <- "/Library/Frameworks/R.framework/Resources/library"
str_libraries <- c("readxl", "phyloseq", "tidyverse", "pacman", "yaml", "ggplot2", "vegan", "microbiome", "ggpubr", "viridis", "decontam", "gridExtra", "ggpubr", "lme4", "lmerTest", "writexl", "harrietr", "Maaslin2", "ggtext", "ggpmisc", "gridExtra", "gamm4", "reshape2")
        
YAML_header <-
'---
title: "Host-DNA depletion 1: data wrangling"
author: "Minsik Kim"
date: "2032.03.10"
output:
    rmdformats::downcute:
        downcute_theme: "chaos"
        code_folding: hide
        fig_width: 6
        fig_height: 6
---'
seed <- "20230310"

#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Loads libraries, file paths, and other document options.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Boot <- function() {
    .libPaths(path_library)

    require(pacman)
    pacman::p_load(c("knitr", "rmarkdown", "rmdformats", "yaml"))

    knitr::opts_knit$set(root.dir = path_working)

    str_libraries |> unique() |> sort() -> str_libraries
    pacman::p_load(char = str_libraries)

    set.seed(seed)
}
FUN.LineZero.Boot()
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#Outputs R environment report.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
FUN.LineZero.Report <- function() {
    cat("Line Zero Environment:\n\n")
    paste("R:", pacman::p_version(), "\n") |> cat()
    cat("Libraries:\n")
    for (str_libraries in str_libraries) {
        paste(
            "    ", str_libraries, ": ", pacman::p_version(package = str_libraries),
            "\n", sep = ""
        ) |> cat()
    }
    paste("\nOperating System:", pacman::p_detectOS(), "\n") |> cat()
    paste("    Library Path:", path_library, "\n") |> cat()
    paste("    Working Path:", path_working, "\n") |> cat()
    paste("Seed:", seed, "\n\n") |> cat()
    cat("YAML Header:\n")
    cat(YAML_header)
}
FUN.LineZero.Report()


```

#Data cleaning description#

1. Loading data 
2. Merging all the output
3. Tidifying metadata
4. Save RDS file


***1. Loading data***


```{r warning=F, `Loading data`}
#set working directory
setwd("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/2_Protocols/host_depletion/")

#Loading meta data
        #Loading the data
        sample_data1 <- read_excel("SOP_20220324_MGK_Host_extraction.xlsx", sheet = 2)
        #Cryopreserved data
        sample_data2 <- read_excel("SOP_20220404_MGK_Host_qPCR_cryopreserve.xlsx", sheet = 2)
        #BAL and sputum data
        sample_data3 <- read_excel("SOP_20220407_MGK_sputum_BAL_host_depletion.xlsx", sheet = 5)
        #Additional BAl and NS data
        sample_data4 <- read_excel("SOP_20220606_MGK_BAL_NB_host_depletion.xlsx", sheet = 2)
        #control_data
        sample_data5 <- read_excel("SOP_20230112_MGK_host_depletion_controls.xlsx", sheet = 2) %>% subset(., .$sample_type != "nasal_swab")
        
#Laoding the qPCR data
        #Loading data  - host
        qPCR_data_host <- function(data){
                data %>% data.frame() %>% mutate(Quantity.Mean = case_when(
                        is.na(Quantity.Mean) ~ Quantity,
                        TRUE ~ Quantity.Mean)) %>%
                rename(extraction_id = "Sample", DNA_host_well = Quantity.Mean) %>% select(c("extraction_id", "DNA_host_well"))
        }
        host_simple1 <- read_excel("SOP_20220324_MGK_Host_extraction.xlsx", sheet = 6) %>% qPCR_data_host()
        host_simple2 <- read_excel("SOP_20220404_MGK_Host_qPCR_cryopreserve.xlsx", sheet = 5) %>% qPCR_data_host()
        host_simple3 <- read_excel("SOP_20220407_MGK_sputum_BAL_host_depletion.xlsx", sheet = 6) %>% qPCR_data_host()
        host_simple4 <- read_excel("SOP_20220606_MGK_BAL_NB_host_depletion.xlsx", sheet = 5) %>% qPCR_data_host()
        host_simple5 <- read_excel("SOP_20230112_MGK_host_depletion_controls.xlsx", sheet = 9) %>% qPCR_data_host()

#Loading data  - Bac
        qPCR_data_bac <- function(data){
                data %>% data.frame() %>% mutate(Quantity.Mean = case_when(
                        is.na(Quantity.Mean) ~ Quantity,
                        TRUE ~ Quantity.Mean)) %>%
                rename(extraction_id = "Sample", DNA_bac_well = Quantity.Mean) %>% select(c("extraction_id", "DNA_bac_well"))
        }
        bac_simple1 <- read_excel("SOP_20220324_MGK_Host_extraction.xlsx", sheet = 7) %>% qPCR_data_bac()
        bac_simple2 <- read_excel("SOP_20220404_MGK_Host_qPCR_cryopreserve.xlsx", sheet = 6) %>% qPCR_data_bac()
        bac_simple3 <- read_excel("SOP_20220407_MGK_sputum_BAL_host_depletion.xlsx", sheet = 7) %>% qPCR_data_bac()
        bac_simple4 <- read_excel("SOP_20220606_MGK_BAL_NB_host_depletion.xlsx", sheet = 6) %>% qPCR_data_bac()
        bac_simple5 <- read_excel("SOP_20230112_MGK_host_depletion_controls.xlsx", sheet = 8) %>% qPCR_data_bac()

```


***Merging all the output***


```{r, warning=F, `Merging all the outputs`}
        #Mergineg sampledata, host and bacterial qPCR data
        sample_results <- list()
        for (i in 1:5) {
                  sample_data <- get(paste0("sample_data", i))
                  host_simple <- get(paste0("host_simple", i))
                  bac_simple <- get(paste0("bac_simple", i))
                  
                  sample_results[[i]] <- merge(sample_data, host_simple, by = "extraction_id") %>%
                                          merge(., bac_simple, by = "extraction_id") %>%
                                          subset(., !duplicated(extraction_id, fromLast = T)) %>%
                          mutate(collection_date = as.character(collection_date), extraction_date = as.character(extraction_date))
        }

        #Changing type of data
        sample_results[[3]]$aliquot <- as.character(sample_results[[3]]$aliquot)
        
        sample_result <- dplyr::bind_rows(sample_results[[1]], sample_results[[2]], sample_results[[3]], sample_results[[4]], sample_results[[5]])

```

***Tidyfying data***

1. Re-calculattion of DNA concentration (adjusting dilution factors, standard concentration, etc.)
2. creating new column (treatment)


```{r warning=F, `Caculation of DNA, assigning treatment groups`}

#Specification at Zymo was 20, 2, 0.2, ... per well. 2 uL of standards were used.
#qPCR standards were 10, 0.1, 0.01 ng / uL etc. 
#Prior calculatoin was based on zymo's specification. (20 ng / uL, which is wrong). This need to be modified

sample_result<- sample_result %>% 
        mutate(DNA_bac_ng_uL = DNA_host_well/2, DNA_host_ng_uL = DNA_bac_well/2,
               DNA_host_nondil = DNA_host_ng_uL * 10, DNA_bac_nondil = DNA_bac_ng_uL * 10, #qPCR was conducted for 10-fold diluted samples.
               host_proportion = (DNA_host_nondil/(DNA_host_nondil + DNA_bac_nondil)))

#Method and sample_type labels modification
sample_result <- sample_result %>% mutate(
        treatment = case_when(
                sample_result$control == 1 ~ "control",
                sample_result$lypma == 1 ~ "lyPMA",
                sample_result$benzonase == 1 ~ "benzonase",
                sample_result$qiaamp == 1 ~ "qiaamp",
                sample_result$host_zero == 1 ~ "host_zero",
                sample_result$molysis == 1 ~ "molysis",
                TRUE ~ "control"),
        sample_type = case_when(
                sample_type == "host_depletion_negative_control" ~ "neg_depletion",
                sample_type == "host_depletion_positive_control" ~ "pos_depletion",
                TRUE ~ sample_type)
)


```

#Merging all the output#                

1. Loading phyloseq object
2. Adding more data to the phyloseq - sampledata
3. Factorize predictors
4. Saving RDS

***1. Loading phyloseq object***


```{r warning=FALSE, `Loading phyloseq object`}

phyloseq1 <- microbiome::read_biom2phyloseq("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20221102_PQ00331_Host_Depletion/metaphlan3/metaphlan3.allkingdoms.EstCount.biom")
phyloseq2 <- microbiome::read_biom2phyloseq("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20230124_PQ00430_sicas2_nasal_host_depleted/LaiP_WGS3_PQ00430_Deliverables_Without_Spikeins/metaphlan3/metaphlan3.allkingdoms.EstCount.biom")


#Merging

readstat1 <- read.csv("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20221102_PQ00331_Host_Depletion/ReadStats.txt", sep = "\t") %>%
        rename(baylor_other_id = SampleID)
#control data
readstat2 <- read.csv("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20230124_PQ00430_sicas2_nasal_host_depleted/LaiP_WGS3_PQ00430_Deliverables_Without_Spikeins/ReadStats.txt", sep = "\t") %>%
        rename(baylor_id = SampleID)

baylor_ids <- rbind(read_excel("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/3_Documentation/Communications/2022-06-22_baylor_shipping_host_depletion_lowinput/CMMR_MetadataCapture_20220622_LaiP-PQ00181_Host DNADepletion.xlsx", skip = 27),
                    baylor_ids2 <- read_excel("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed/20230124_PQ00430_sicas2_nasal_host_depleted/CMMR_MetadataCapture_20230124_LaiP-PQ00430_SICAS2_NS.xlsx", skip = 27)
)

baylor_ids <- baylor_ids %>% select(c("UNIQUE alphanumeric ID", "Optional              secondary ID")) %>%
        rename(baylor_id = "UNIQUE alphanumeric ID",
               baylor_other_id = "Optional              secondary ID",) %>%
        subset(., !is.na(.$baylor_other_id))

#Merging
readstat <- rbind(merge(readstat1, baylor_ids, by = "baylor_other_id"),
                  merge(readstat2, baylor_ids, by = "baylor_id"))

#Generating baylor ID
sample_result <- sample_result %>%
        mutate(baylor_other_id = case_when(
            extraction_date == "20230119" ~ sample_id,
            sample_type == "neg_extraction" ~ paste(extraction_date, "Neg", sep = "_"),
            sample_type == "pos_extraction" ~ paste(extraction_date, "Pos", sep = "_"),
            cryo_preserve == 1 ~ paste(sample_id, "cryo", sep = "_"),
            sample_type == "BAL" ~ paste(sample_id, aliquot, treatment, sep = "_"),
            grepl("NS_26|NS_37", sample_id) ~ paste(sample_id, aliquot, treatment, sep = "_"),
            TRUE ~ paste(sample_id, treatment, sep = "_"))) %>%
        mutate(baylor_other_id = gsub("20220610", "20220606", baylor_other_id)) # older extraction date used for baylor shipping log
        

#Merging Readstat + metadata
sample_data1 <- merge(readstat, sample_result, by = "baylor_other_id") %>%
        mutate(baylor_other_id2 = baylor_other_id) %>%
        column_to_rownames(var = "baylor_other_id2") %>%
        sample_data()
sample_data2 <- merge(readstat, sample_result, by = "baylor_other_id") %>%
        mutate(baylor_id2 = baylor_id) %>%
        column_to_rownames(var = "baylor_id2") %>%
        sample_data()
phyloseq <- merge_phyloseq(merge_phyloseq(sample_data(sample_data1), phyloseq1), merge_phyloseq(sample_data(sample_data2), phyloseq2))



```

***2. Adding more data to the phyloseq - sampledata***

```{r, warning=F, `Adding more data to phyloseq object`}
#Adding host DNA proportion of sequencing results
sample_data <- sample_data(phyloseq) %>% data.frame %>% 
        mutate(sequencing_host_prop = Host_mapped/Reads_after_trim, .after = Metaphlan_mapped)

#Adding Treatment status
sample_data$treated <- rowSums(sample_data[, c("lypma", "benzonase", "molysis", "host_zero", "qiaamp")])
sample_data$treated <- ifelse(is.na(sample_data$treated), 0, sample_data$treated)

# Adding library prep results --------------------------------------------
lib_fail_list <- c("NS_6_b_lyPMA", "NS_8_b_lyPMA", "NS_17_b_lyPMA", "NS_21_b_host_zero", "NS_22_b_host_zero", "NS_19_d_molysis", "NS_21_d_molysis", "NS_22_d_molysis", "NS_23_d_molysis", "NS_26_b_lyPMA", "BAL_073_b_host_zero", "BAL_073_c_molysis", "BAL_078_d_lyPMA")
sample_data$lib_failed <- sample_data$baylor_other_id %in% lib_fail_list



```


***2. Adding more data to the phyloseq - picogreen***

```{r, warning=F, `Adding more data to phyloseq object`}
#Adding host DNA proportion of sequencing results
picogreen <- rbind(read_excel("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/DNA_Concentration/DAR_20221221_MGK_picogreen_PQ00331.xlsx", sheet = 1, skip = 1),
       read_excel("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/DNA_Concentration/DAR_20221221_MGK_picogreen_PQ00331.xlsx", sheet = 2, skip = 1)
       ) %>% rename(baylor_other_id = "Sample", picogreen_ng_ul = "Concentration (ng/ul)") %>% select(c("baylor_other_id", 
                                                                                                        "picogreen_ng_ul"))
sample_data <- merge(sample_data %>% rownames_to_column("row.names"), picogreen, all = T, by = "baylor_other_id") %>% subset(.,!is.na(.$row.names)) %>% remove_rownames() %>% column_to_rownames("row.names")


#constructing phylose object
phyloseq <- merge_phyloseq(otu_table(phyloseq), tax_table(phyloseq), sample_data(sample_data))



```




***3. Factorize predictors***

```{r, warning=F, `Factorize predictors`}
# Factorize predicters ----------------------------------------------------
# + creating lables for figures

sample_data(phyloseq) <- sample_data(phyloseq) %>% data.frame(check.names = F) %>% 
        mutate(treatment = tolower(treatment), 
               treatment = factor(treatment,
                                  levels = c("control", "lypma", "benzonase", "host_zero", "molysis", "qiaamp"),
                                  labels = c("Untreated", "lyPMA", "Benzonase", "Host zero", "Molysis", "QIAamp")),
               sample_type = factor(sample_type,
                                    levels = c("neg_depletion", "neg_extraction", "pos_extraction", "pos_depletion", "BAL", "nasal_swab", "Sputum"), 
                                    #labels = c("BAL", "Nasal", "Sputum", "Neg. ext.", "Pos. ext.", "Neg. dep.", "Pos. dep."))) %>%
                                    labels = c( "Neg.", "Neg.", "Mock", "Mock", "BAL", "Nasal", "Sputum"))) %>%
        #rename(subject_id = "original_sample") %>%
        sample_data()


sample_data(phyloseq)$subject_id <- ifelse(is.na(sample_data(phyloseq)$subject_id), as.character(sample_data(phyloseq)$sample_type), sample_data(phyloseq)$subject_id)
sample_data(phyloseq)$subject_id <- ifelse(sample_data(phyloseq)$subject_id == "lab_member", as.character(sample_data(phyloseq)$original_sample),
                                           sample_data(phyloseq)$subject_id)



```

Adding functional data



```{r warning=FALSE, `Loading phyloseq object`}
                 
#Loading phyloseq object
setwd("/Users/minsikkim/Dropbox (Partners HealthCare)/Project_Baylor/4_Data/1_Raw/Baylor_Processed")
path_rpk <- read.csv("20221102_PQ00331_Host_Depletion/humann3/PathAbundance.relab.metagenome.txt", sep = "\t", check.names = F, row.names = "# Pathway") %>% select(-c("taxonomy"))
path_tax <- read.csv("20221102_PQ00331_Host_Depletion/humann3/PathAbundance.relab.metagenome.txt", sep = "\t", check.names = F) %>%
        select(c("taxonomy", "# Pathway")) %>% rename(pathway = "# Pathway", group = "taxonomy")
row.names(path_tax) <- path_tax$pathway

path_rpk2 <- read.csv("20230124_PQ00430_sicas2_nasal_host_depleted/LaiP_WGS3_PQ00430_Deliverables_Without_Spikeins/humann3/PathAbundance.RPK.metagenome.txt", sep = "\t", check.names = F, row.names = "# Pathway") %>% select(-c("taxonomy"))
path_tax2 <- read.csv("20230124_PQ00430_sicas2_nasal_host_depleted/LaiP_WGS3_PQ00430_Deliverables_Without_Spikeins/humann3/PathAbundance.RPK.metagenome.txt", sep = "\t", check.names = F) %>%
        select(c("taxonomy", "# Pathway")) %>% rename(pathway = "# Pathway", group = "taxonomy")
row.names(path_tax2) <- path_tax2$pathway

#Merging Readstat + metadata
phyloseq_path_rpkm1 <- #merge_phyloseq(
        merge_phyloseq(otu_table(path_rpk, taxa_are_rows = T), sample_data(phyloseq), tax_table(path_tax %>% as.matrix()))#,
        #merge_phyloseq(otu_table(path_rpk2, taxa_are_rows = T), sample_data(phyloseq), tax_table(path_tax2 %>% as.matrix()))

phyloseq_path_rpkm2 <- #merge_phyloseq(
        merge_phyloseq(otu_table(path_rpk2, taxa_are_rows = T), sample_data(phyloseq), tax_table(path_tax2 %>% as.matrix()))#,
        #merge_phyloseq(otu_table(path_rpk2, taxa_are_rows = T), sample_data(phyloseq), tax_table(path_tax2 %>% as.matrix()))
phyloseq_path_rpkm <- merge_phyloseq(phyloseq_path_rpkm1, phyloseq_path_rpkm2)

```

```{r warning=FALSE, `Loading functional - RPKM`}

phy <- list()
sample_data <- sample_data(phyloseq)
sample_names(phyloseq) <- sample_data$baylor_other_id
sample_names(phyloseq_path_rpkm) <- sample_data$baylor_other_id
phy$phyloseq_count <- phyloseq
phy$phyloseq_rel <- transform_sample_counts(phyloseq, function(x){x / sum(x)})
phy$phyloseq_path_rpkm <-phyloseq_path_rpkm


tax_table(phy$phyloseq_count) <- cbind(tax_table(phy$phyloseq_count),
                                       AMR::mo_gramstain(taxa_names(phy$phyloseq_count))) %>%
        tax_table()
colnames(tax_table(phy$phyloseq_count))[8] <- "Gram"

tax_table(phy$phyloseq_rel) <- cbind(tax_table(phy$phyloseq_rel),
                                       AMR::mo_gramstain(taxa_names(phy$phyloseq_rel))) %>%
        tax_table()
colnames(tax_table(phy$phyloseq_rel))[8] <- "Gram"
AMR::mo_gramstain(taxa_names(phy$phyloseq_count))

sample_data(phy$phyloseq_rel)$gram_neg <- phy$phyloseq_rel %>% subset_taxa(Gram == "Gram-negative") %>% otu_table() %>% colSums(na.rm = F)
sample_data(phy$phyloseq_count)$gram_neg <- phy$phyloseq_rel %>% subset_taxa(Gram == "Gram-negative") %>% otu_table() %>% colSums(na.rm = F)
sample_data(phy$phyloseq_path_rpkm)$gram_neg <- phy$phyloseq_rel %>% subset_taxa(Gram == "Gram-negative") %>% otu_table() %>% colSums(na.rm = F)
```



***4. Saving RDS***

```{r, warning=F, `Saving phyloseq object`}
# Saving object -----------------------------------------------------------
#write.csv(sample_result, "/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/7_Manuscripts/2022_MGK_Host_Depletion/Tables/DAT_20230122_MGK_host_control_qPCR.csv")
write_rds(phy, "/Users/minsikkim/Dropbox (Partners HealthCare)/Project_SICAS2_microbiome/4_Data/2_Tidy/Phyloseq/PHY_20221129_MGK_host_tidy_tax.rds")


```



```{r warning=FALSE }
#===============================================================================
#BTC.LineZero.Footer.1.1.0
#===============================================================================
#R markdown citation generator.
#===============================================================================
#RLB.Dependencies:
#   magrittr, pacman, stringr
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
#BTC.Dependencies:
#   LineZero.Header
#===============================================================================
#Generates citations for each explicitly loaded library.
#=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
str_libraries <- c("r", str_libraries)
for (str_libraries in str_libraries) {
    str_libraries |>
        pacman::p_citation() |>
        print(bibtex = FALSE) |>
        capture.output() %>%
        .[-1:-3] %>% .[. != ""] |>
        stringr::str_squish() |>
        stringr::str_replace("_", "") |>
        cat()
    cat("\n")
}
#===============================================================================
```